A Classical Density-Functional Theory for Describing Water Interfaces

### 2013-08-23 papers/water-saft/paper.tex

We will apply this improved $g_\sigma^{HS}$ to a SAFT-based
classical density functional for water~\cite{hughes2013classical}.
This functional was constructed so as to reproduce in the homogeneous
limit the 4-site optimal SAFT model for water developed by Clark
\emph{et al.}~\cite{clark2006developing}.  The DFT of
Ref.~\citenum{hughes2013classical} uses the association free energy
functional of Yu and
Wu~\cite{yu2002fmt-dft-inhomogeneous-associating}, which is based on a
$g_\sigma^{HS}$ that we have since found to be
inaccurate~\cite{schulte2012using}.  In this paper, we will examine
the result of using the improved functional for $g_\sigma^{HS}$
developed in Ref.~\citenum{schulte2012using} to construct an
association free energy functional.

@@

\begin{figure}
\begin{center}
\includegraphics[width=3.5in]{figs/density-compare}
\end{center}
\caption{ Density profiles for a water around a single hydrophobic
  rod of radius 0.2 nm. The solid red profile is the density using our new functional
  and the dashed blue profile the result of our old functional.  For
  scale, under the profiles is a cartoon of a string of hard
  spheres touching in one dimension. The horizontal black dotted line
  is the bulk density for water.}
\label{fig:density-single-rod}
\end{figure}

@@

We will first discuss the case of a single hydrophobic rod immersed in
water. Figure~\ref{fig:density-single-rod} shows the density profile
of water near a rod with radius~0.1~\AA\fixme{Eric, please figure this
  radius out}.  The density computed using the functional of this
paper is qualitatively similar to that from
Ref.~\citenum{hughes2013classical}, with a comparable density at
contact---consistent with having made only a moderate change in the
free energy.  The first density peak near the surface is consistently
higher than that from Ref.~\citenum{hughes2013classical}, and the peak
is not smooth.  This results from the improved accuracy of the
$g_\sigma^\textit{HS}$ from Ref.~\citenum{schulte2012using}, since beyond the
first peak water molecules are unable to touch---or hydrogen bond
to---molecules at the surface of the hard rod.

\begin{figure}
\begin{center}
\includegraphics[width=3.5in]{figs/single-rod-broken-HB}
\end{center}
\caption{Broken hydrogen bonds per nanometer for various hydrophobic rods
  immeresed in water.  The solid red line uses our new functional
  while the dashed blue line uses the old one. For large enough rods,
  the graph increases linearly for both functionals.}
\label{fig:single-rod-broken-HB}
\end{figure}

@@

\begin{figure}
\begin{center}
\includegraphics[width=3.5in]{figs/sphere-broken-HB}
\end{center}
\caption{Broken hydrogen bonds per for various hard spheres
  immeresed in water.  The solid red line uses our new functional
  while the dashed blue line uses the old one.}
\label{fig:spheres-broken-HB}
\end{figure}

>>>>

We will apply this improved $g_\sigma^{HS}$ to a SAFT-based classical
density functional for water~\cite{hughes2013classical}.  This
functional was constructed to reduce in the homogeneous limit
to the 4-site optimal SAFT model for water developed by Clark \emph{et
  al.}~\cite{clark2006developing}.  The DFT of
Ref.~\citenum{hughes2013classical} uses the association free energy
functional of Yu and
Wu~\cite{yu2002fmt-dft-inhomogeneous-associating}, which is based on a
$g_\sigma^{HS}$ that we have since found to be
inaccurate~\cite{schulte2012using}.  In this paper, we will examine
the result of using the improved functional for $g_\sigma^{HS}$
developed in Ref.~\citenum{schulte2012using} to construct an
association free energy functional.

@@

\begin{figure}
\begin{center}
\includegraphics[width=3.5in]{figs/density-compare}
\end{center}
\caption{ Density profiles for a water around a single hydrophobic rod
  of radius 0.1 nm. The solid red profile is from the functional
  developed in this paper and the dashed blue profile is the result
  from Ref. \cite{hughes2013classical}.  For scale, under the profiles
  is a cartoon of a string of hard spheres touching in one
  dimension. The horizontal black dotted line is the bulk density for
  water and the vertical line on the left represents the wall of the rod.}
\label{fig:density-single-rod}
\end{figure}

@@

We will first discuss the case of a single hydrophobic rod immersed in
water. Figure~\ref{fig:density-single-rod} shows the density profile
of water near a rod with radius~1~\AA.  The density computed using the functional of this
paper is qualitatively similar to that from
Ref.~\citenum{hughes2013classical}, with a comparable density at
contact---consistent with having made only a moderate change in the
free energy.  The first density peak near the surface is consistently
higher than that from Ref.~\citenum{hughes2013classical}, and the peak
is not smooth.  This results from the improved accuracy of the
$g_\sigma^\textit{HS}$ from Ref.~\citenum{schulte2012using}, since beyond the
first peak water molecules are unable to touch---or hydrogen bond
to---molecules at the surface of the hard rod.

\begin{figure}
\begin{center}
\includegraphics[width=3.5in]{figs/single-rod-broken-HB}
\end{center}
\caption{Broken hydrogen bonds per nanometer for hydrophobic rods
  immeresed in water.  The solid red line uses the functional developd
  in this paper while the dashed blue line uses the functional from
  Ref. \cite{hughes2013classical}. For large enough rods, the graph
  increases linearly for both functionals.}
\label{fig:single-rod-broken-HB}
\end{figure}

@@

\begin{figure}
\begin{center}
\includegraphics[width=3.5in]{figs/sphere-broken-HB}
\end{center}
\caption{Broken hydrogen bonds for hard spheres immeresed in water.
  The solid red line uses our the functional developd in this paper
  while the dashed blue line is from
  Ref. \cite{hughes2013classical}l.}
\label{fig:spheres-broken-HB}
\end{figure}
